---
title: "Supplement to 'The First Australian Plant Foods at Madjedbebe, 65,000–53,000 years ago'"
author:
  - Ben Marwick
  - S. Anna Florin
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
---


<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

This repository includes code and data for Supplementary Figure 1

```{r}
# originally from analysis done in preparation for our response to comment from Kaifu et al.
library(readr)
library(viridis)
library(here)

# c14_ages and osl_ages are produced by data and code in https://github.com/benmarwick/mjbnaturepaper

c14_ages  <- read_csv(here("analysis/data/raw_data/c14_ages.csv"))
osl_ages  <- read_csv(here("analysis/data/raw_data/osl_ages.csv"))


# prepare data
excluding <- c("Wk43605" , # pit feature
               #"Wk43606", 
               #"Wk43607", 
               #"Wk43610", 
               #"Wk43604",
               #"Wk43611",
               #"OZT591",
               "OZT593"
)

# excluding <- NA

squares <- "B3|B4|B5|C3|C4|C5|D3|D4|D5|E3|E4|E5"

c14_ages_excludes <-  c14_ages[!grepl(paste0(excluding, collapse = "|"), c14_ages$Lab.ID), ]
c14_ages_excludes <- c14_ages_excludes[grepl(paste0(squares, collapse = "|"), c14_ages_excludes$square), ]

osl_ages_NWSWABC <-  osl_ages[grepl("NW.|SW.*A|SW.*B|SW.*C", osl_ages$Sample), ]
osl_ages_NE <- osl_ages[grepl("NE.", osl_ages$Sample), ]
```


```{r}
# test for difference in age-depth models

# combine data with grouping variable
ages_with_groups <- data.frame(ages = c(c14_ages_excludes$Bchron_Median/1000, 
                                        osl_ages_NWSWABC$osl_age),
                               depths = c(c14_ages_excludes$depth_below_surface,
                                          osl_ages_NWSWABC$total_station_depth_below_surf),
                               group = c(rep("c14", nrow(c14_ages_excludes)),
                                         rep("osl NWSWABC", nrow(osl_ages_NWSWABC))))

#model without grouping variable
fit0 <- lm(ages ~ poly(depths, 2), data = ages_with_groups)

#model with grouping variable
fit1 <- lm(ages ~ poly(depths, 2) * group, data = ages_with_groups)

#compare models 
all_osl_NWSWABC_c14_BCDE_3_5 <- anova(fit0, fit1)
pval <- round(all_osl_NWSWABC_c14_BCDE_3_5$`Pr(>F)`[2], 4)
fval <- round(all_osl_NWSWABC_c14_BCDE_3_5$F[2], 3)
df <- all_osl_NWSWABC_c14_BCDE_3_5$Df[2]
```


```{r}
library(ggrepel)

formula <- y ~ poly(x, 2)

p <- 
  ggplot() +
  geom_point(data = osl_ages_NE,
             aes(-total_station_depth_below_surf, 
                 osl_age),
             colour = viridis(3)[2]) +
  geom_text_repel(data = osl_ages_NE,
                  aes(-total_station_depth_below_surf, 
                      osl_age,
                      label = Sample),
                  size = 1.5) +
  geom_point(data = osl_ages_NWSWABC,
             aes(-total_station_depth_below_surf, 
                 osl_age),
             colour = viridis(3)[1]) +
  geom_smooth(data = osl_ages_NWSWABC,
              aes(-total_station_depth_below_surf, 
                  osl_age),
              method = "lm", 
              formula = formula, 
              #fullrange = TRUE,
              colour = viridis(3)[1]) +
  geom_text_repel(data = osl_ages_NWSWABC,
                  aes(-total_station_depth_below_surf, 
                      osl_age,
                      label = Sample),
                  size = 1.5) +
  geom_point(data =c14_ages_excludes,
             aes(-depth_below_surface, 
                 Bchron_Median/1000),
             colour = viridis(3)[3]) +
  geom_smooth(data = c14_ages_excludes,
              aes(-depth_below_surface, 
                  Bchron_Median/1000),
              method = "lm", 
              formula = formula, 
              # fullrange = TRUE,
              colour = viridis(3)[3]) +
  geom_text_repel(data = c14_ages_excludes,
                  aes(-depth_below_surface, 
                      Bchron_Median/1000,
                      label = Lab.ID),
                  size = 1.5)  +
  theme_minimal(base_size = 14) +
  ylab("Age (kya)") +
  xlab("Depth (m)") 

p 

ggsave(here("analysis/figures/OSL ages from NW, SW A-B-C and C14 ages from B-C-D-E 3-5 no-text.png"), dpi = 600)
ggsave(here("analysis/figures/OSL ages from NW, SW A-B-C and C14 ages from B-C-D-E 3-5 no-text.svg"))
```


Age depth curve model, comparing OSL ages (in purple) from the northwest and southwest sides of Squares A, B and C to AMS radiocarbon ages (in yellow) from Squares B, C, D and E, 3-5 (excluding WK43606 and OZT593, see Supplementary Fig. 2c). Green points show OSL ages from the NE side for reference, these are not included in the statistical models. The age-depth models for C14 ages and OSL ages are not significantly different (ANOVA on linear models with quadratic terms, `r paste0("F = ", fval, ", df = ", df, ", p = ", pval)`)

<!-- The following line inserts a page break  -->
\newpage

```{r}
p_plus_text <- 
  p +
  annotate("text", x = 2.7, y = 10, 
           label = paste0("Excluding:\n", paste0(excluding, collapse = "\n")), 
           size = 2)  +
  annotate("text",  x = 0.5, y = 75, hjust = 0,
           label = paste0("Age-depth model comparison ANOVA\n",
                          "Comparing OSL ages from NW & SW A-B-C to\n",
                          "C14 ages from B-C-D-E 3-5:\n",
                          "F = ", fval, ", df = ", df, ", p = ", pval, " (quadratic fits)",
                          "\nModels are ", ifelse(pval <= 0.05, "", "not "), "significantly different" ), 
           size = 3) +
  ggtitle("OSL ages and C14 ages from B-C-D-E 3-5") +
    theme_minimal(base_size = 10) 

p_plus_text

ggsave(here("analysis/figures/OSL ages from NW, SW A-B-C and C14 ages from B-C-D-E 3-5.png"), dpi = 600)
ggsave(here("analysis/figures/OSL ages from NW, SW A-B-C and C14 ages from B-C-D-E 3-5.svg"))
```




\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies: 

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```
